Binocular integration of chromatic and luminance signals

### 2024-03-24 manuscript.tex
Fixed Stan model!

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage{float} \floatplacement{figure}{H} \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{A\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{A\arabic{figure}}}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Binocular integration of chromatic and luminance signals},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

@@

The process by which the brain combines independent inputs is of fundamental importance for understanding sensory perception. Binocular vision is a useful test-case for determining the general principles involved in neural signal combination, as our brains combine the inputs from the left and right eyes to provide binocular single vision. In recent years our understanding has been facilitated by the development of binocular gain control models that provide a framework to interpret empirical data across multiple techniques, including psychophysics (Meese et al., 2006), EEG (Baker \& Wade, 2017), fMRI (Moradi \& Heeger, 2009) and pupillometry (Segala et al., 2023). However, the majority of this work has used achromatic (black and white) stimuli; we know comparatively little about how chromatic signals are combined binocularly, or about how signals in different ocular and chromatic channels interact. In this study we use psychophysical detection and discrimination paradigms to explore binocular interactions in the chromatic pathways.

@@

The detailed pattern of results across these four conditions is complex, and for achromatic stimuli has several distinctive features that have been replicated in multiple studies. At low pedestal contrasts, the binocular condition yields lower thresholds than the monocular condition owing to binocular summation (Baker et al., 2018; Campbell \& Green, 1965). However at high pedestal contrasts the `handle' regions of the dipper functions for these conditions converge: a consequence of interocular suppression compensating for the increased excitation during binocular stimulation (Legge, 1984; Maehara \& Goryo, 2005). The half-binocular condition avoids confounding the number of eyes seeing the target with the number of eyes seeing the pedestal (Meese et al., 2006). The pedestal is always binocular in this condition, whereas the target increment is monocular, and thresholds are consistently higher than in the binocular condition across the full range of pedestal contrasts. This demonstrates that binocular summation occurs across the full contrast range, when the pedestal ocularity is appropriately controlled. Finally, the dichoptic condition produces extremely strong masking of the target, such that when the pedestal is visible, the target must equal or exceed its contrast in order to be detectable (Baker \& Meese, 2007; Legge, 1979; Maehara \& Goryo, 2005).

At the output of the human retina, visual signals are split into three chromatic pathways. The sum of long- and medium-wavelength cone outputs (L+M) transmits luminance information, and is likely responsible for the binocular combination effects previously studied using achromatic stimuli (see above). The difference of long- and medium-wavelength cone outputs (L-M) is responsive to chromatic stimuli modulating along a red/green axis in colour space. Finally the short wavelength cone outputs (S-(L+M)) code chromatic stimuli modulating along a blue/yellow axis. There has not yet been a detailed investigation of binocular contrast discrimination in either of these chromatic pathways, however there is reason to believe they may differ from the achromatic pathway. At detection threshold, binocular summation is greater for chromatic versus achromatic stimuli (Simmons, 2005), implying a more linear initial stage of processing. For cross-orientation masking, there are differences in the magnitude of masking between chromatic (red/green) and achromatic stimuli (Kim et al., 2013; Medina \& Mullen, 2009), as well as differences in their temporal dynamics (Kim \& Mullen, 2015). There are also interactions between chromatic and achromatic pathways both within (Chen et al., 2000) and between (Kingdom \& Libenson, 2015; Mullen et al., 2014) the eyes, yet these have not been fully explored for arrangements where the target and mask have the same orientation. Finally, the neurophysiological underpinnings of colour vision are distinct from the achromatic system. In primary visual cortex (V1), chromatic signals are processed in `blob' regions that are revealed by cytochrome oxidase staining (Horton \& Hubel, 1981). The blob regions appear to be largely monocular (Livingstone \& Hubel, 1984), suggesting that binocular combination for chromatic stimuli might occur later than for achromatic stimuli and be subject to different constraints.

@@

All experiments took place in a darkened room. Participants placed their heads in a chin rest mounted on a height-adjustable table, to which the stereoscope was also attached. The total optical viewing distance (including the light path through the mirrors) was XXcm, at which distance 1 degree of visual angle encompassed 48 pixels on the monitor.

Before beginning primary data collection, each participant in Experiments 1 and 2 completed an isoluminance adjustment task. Stimuli were presented that counterphase flickered at 5Hz, defined about either the L-M or S-(L+M) plane in cone space. Participants used a trackball to dynamically adjust the angle of the stimulus to minimise the percept of flicker. Each participant completed ten such trials for each colour plane, and the average angle across repetition was taken as the isoluminant point, and used to generate stimuli for the main experiment for that participant. Settings were very similar across participants for the S-(L+M) direction, and somewhat more heterogeneous for the L-M direction (see Figure \ref{fig:isofig}).

@@

In Experiment 1, participants completed a two-interval-forced-choice (2IFC) contrast discrimination task. Stimuli were presented for 200ms, with an interstimulus interval of 400ms. Each interval was indicated by an auditory beep, and participants made their responses using a two-button trackball. Correct trials were indicated by a high pitched tone, and incorrect trials by a low pitched tone. Each block of the experiment tested a single pedestal contrast level, and lasted around 12 minutes. On each trial the target contrast level was determined by a 3-down-1-up staircase procedure. There were 8 interleaved staircases in total; four stimulus arrangements (see Figure \ref{fig:exampledips}) by two target eye assignments. Each pedestal contrast was repeated 3 times by each participant, and the block order was randomised. The experiment lasted around 4 hours per participant for each chromatic condition, which took place over the course of several weeks. In total, the experiment consisted of 67978 trials (pooled across participants).

@@

In Experiment 3, the achromatic conditions from Experiment 1 were repeated using a flickering disc stimulus. The stimulus counterphase flickered at 4Hz, and was presented for 500ms (i.e.~2 full cycles of the temporal modulation). All other procedures were the same as for Experiment 1, and the experiment comprised a total of X trials (pooled across participants).

@@

\(Stage1_L = \frac{C_L^m}{S + C_L + \omega C_R}\)

\(Stage1_R = \frac{C_L^m}{S + C_R + \omega C_L}\)

\(binsum = Stage1_L + Stage1_R\)

\(binresp = \frac{binsum^p}{Z + binsum^q}\)

where \(C_L\) and \(C_R\) are the contrasts displayed to the left and right eyes, and \(m\), \(S\), \(\omega\), \(p\), \(q\) and \(Z\) are free parameters in the model. A further free parameter, \(k\), represents additive internal noise, and is used to convert the model outputs to either d-prime or threshold values. Thresholds are defined when the following equality is satisfied:

\(binresp_{target+pedestal} - binresp_{pedestal} = k\)

and d-prime is defined as:

\(d' = \frac{binresp_{target+pedestal} - binresp_{pedestal}}{k}\).

We also implemented a Bayesian hierarchical version of the model using the Stan language (Carpenter et al., 2017). This used a binomial generator function to model the proportion correct data at each target level, and was fit simultaneously to all participants and each chromatic condition of Experiment 1. This modelling primarily focuses on examining posterior parameter distributions, rather than a model comparison approach. We generated over 1 million posterior samples for the model, and retained 10\% of them for plotting.

@@

All experimental code, raw data and analysis scripts are available at: \url{https://osf.io/3vdga/}. The linked GitHub repository also contains a fully reproducible version of the manuscript. Note that we deviated slightly from the planned preregistration, in that we did not collect data for chromatic flickering discs, or for the cross-pathway dichoptic experiment using disc stimuli. This is because the grating data from Experiments 1 and 2, and the achromatic disc data from Experiment 3, were sufficient to address the questions we had wanted to answer from these experiments.

@@

Dipper functions from Experiment 1 are displayed in Figure \ref{fig:dipperfig}. Panel (a) shows the achromatic results, which replicate the key features from previous work. At detection threshold, binocular summation was a factor of 1.67 (4.47dB), consistent with previous reports. Pedestal masking functions follow the typical `dipper' shape in all conditions, with a region of facilitation at low pedestal contrasts, and masking at higher contrasts. The monocular and binocular dipper handles converge at high contrasts, whereas the half-binocular thresholds remain above the binocular thresholds across the full range of pedestal contrasts. The dichoptic condition produced very high thresholds, with the rising portion of the dipper having a slope around 1 (regression slope of 1.06 across the highest 4 pedestal contrasts in log (dB) units).

@@

\caption{Dipper functions from Experiment 1, averaged across three participants. Error bars show ±1SE across participants. Note that contrast values are expressed as a percentage of the maximum displayable contrast (see Procedures for details). Curves show...}\label{fig:dipperfig}
\end{figure}

A similar pattern of results was observed for both the red/green and blue/yellow isoluminant stimuli (see Figure \ref{fig:dipperfig}b,c). Summation at threshold was a factor of 1.71 (4.66dB) for the red/green targets, and a factor of 1.73 (4.77dB) for the blue/yellow targets, and so was marginally higher than for achromatic stimuli. The general character of the dipper functions was largely consistent with the achromatic results, though we observed shallower facilitation and weaker masking, especially for the blue/yellow stimuli. For example, the strongest facilitation in the binocular condition for achromatic stimuli was a factor of 2.63, whereas it reduced to a factor of 2.35 for red/green stimuli and 1.57 for blue/yellow stimuli. The slope of the binocular dipper handle was 0.52 for achromatic stimuli, 0.58 for red/green stimuli, and 0.34 for blue/yellow stimuli. Dichoptic masking remained as strong for the chromatic conditions as for the achromatic stimuli (regression slopes of 1.3 for red/green and 1.21 for blue/yellow). The pattern of results for individual participants was consistent with the averages, as shown in Figure \ref{fig:individualdips}.

For consistency with previous work, we initially performed least-squares fits of the two-stage model (Meese et al., 2006) for each of the chromaticity experiments (7 free parameters per fit). The best model fits are shown in Figure \ref{fig:dipperfig}, and provide an excellent description of the data, with RMS errors between 1.37 and 1.59dB. We additionally implemented a hierarchical Bayesian version of the model in order to estimate full posterior parameter distributions. This model was fit simultaneously to the full trial-by-trial data from all participants and chromaticity experiments. Figure \ref{fig:bayesianmodel}a summarises the model behaviour, which displays the same pattern of dipper functions as we found empirically. Figure \ref{fig:bayesianmodel}b-h shows posterior distributions for each model parameter. Note in particular that the weight of interocular suppression overlaps 1 for each chromaticity experiment, consistent with the strong dichoptic masking observed in the threshold data.

\caption{Model parameters}\label{fig:bayesianmodel}
\end{figure}

@@

In Experiment 2 we focussed on the dichoptic condition at a single mask contrast, and measured full psychometric functions using the method of constant stimuli for all factorial pairings of target and mask chromaticity. The pooled results across three participants are shown in Figure \ref{fig:MCSfig}, and results for individual participants are available in Figure \ref{fig:individualMCS}. All conditions produced monotonically increasing psychometric functions (panels a-c), but the extent of masking was highly dependent on the relationship between the target and mask chromaticity. Figure \ref{fig:MCSfig}d-f shows a two-dimensional representation of individual threshold and slope estimates (points), as well as the posterior density estimates for fits to the pooled data (ellipses).

@@

We also calculated the slope of the psychometric function for each condition, converted to equivalent Weibull \(\beta\) values. In the absence of a mask, the average slope was \(\beta\) = 3.42, which is typical for contrast detection tasks. Slopes became substantially steeper when the dichoptic mask matched the target in chromaticity (average \(\beta\) = 5.38). These `super-steep' psychometric functions for dichoptic pedestal masking have been reported previously (Baker et al., 2013; Meese et al., 2006), and are observed for the first time here using chromatic stimuli (see diagonal values in Figure \ref{fig:MCSfig}h). However we did not see such markedly steep functions for any of the cross-chromaticity masking conditions (average \(\beta\) = 2.83 for the off-diagonal values).

@@

In our final experiment, we again measured dipper functions, but this time for a temporally modulating luminance disc. This was motivated by our recent work (Segala et al., 2023) that appeared to show increased binocular facilitation and reduced interocular suppression for flickering disc stimuli (relative to gratings), measured using EEG and a psychophysical matching paradigm. The pattern of dipper functions for a 4Hz flickering disc (see Figure X) was very similar to that observed for achromatic gratings (see Figure \ref{fig:dipperfig}a), and the binocular summation ratio at threshold was also similar (a factor of). Fitting the same computational model to the disc data also generated an estimated weight of interocular suppression around 1. Data for individual participants are shown in Figure AX.

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage{float} \floatplacement{figure}{H} \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{A\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{A\arabic{figure}}}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Binocular integration of chromatic and luminance signals},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

@@

The process by which the brain combines independent inputs is of fundamental importance for understanding sensory perception. Binocular vision is a useful test-case for determining the general principles involved in neural signal combination, as our brains typically combine the inputs from the left and right eyes to provide binocular single vision. In recent years our understanding has been facilitated by the development of binocular gain control models that provide a framework to interpret empirical data from multiple experimental paradigms and techniques, including psychophysics (Meese et al., 2006), EEG (Baker \& Wade, 2017), fMRI (Moradi \& Heeger, 2009) and pupillometry (Segala et al., 2023). However, the majority of this work has used achromatic (black and white) stimuli; we know comparatively little about how chromatic signals are combined binocularly, or about how signals in different ocular and chromatic channels interact. In this study we use psychophysical detection and discrimination paradigms to explore binocular interactions in the chromatic pathways.

@@

The detailed pattern of results across these four conditions is complex, and for achromatic stimuli has several distinctive features that have been replicated in multiple studies. At low pedestal contrasts, the binocular condition yields lower thresholds than the monocular condition owing to physiological binocular summation by neurons responsive to signals from both eyes (Baker et al., 2018; Campbell \& Green, 1965). However at high pedestal contrasts the `handle' regions of the dipper functions for these conditions converge: a consequence of interocular suppression compensating for the increased excitation during binocular stimulation (Legge, 1984; Maehara \& Goryo, 2005). The half-binocular condition avoids confounding the number of eyes seeing the target with the number of eyes seeing the pedestal (Meese et al., 2006). The pedestal is always binocular in this condition, whereas the target increment is monocular, and thresholds are consistently higher than in the binocular condition across the full range of pedestal contrasts. This demonstrates that binocular summation occurs across the full contrast range, when the pedestal ocularity is appropriately controlled. Finally, the dichoptic condition produces extremely strong masking of the target, such that when the pedestal is visible, the target must equal or exceed its contrast in order to be detectable (Baker \& Meese, 2007; Legge, 1979; Maehara \& Goryo, 2005).

At the output of the human retina, visual signals are split into three distinct pathways. The sum of long- and medium-wavelength cone outputs (L+M) transmits luminance information, and is likely responsible for the binocular combination effects previously studied using achromatic stimuli (see above). The difference of long- and medium-wavelength cone outputs (L-M) is responsive to chromatic stimuli modulating along a red/green axis in colour space. Finally the short wavelength cone outputs (S-(L+M)) code chromatic stimuli modulating along a blue/yellow axis. There has not yet been a detailed investigation of binocular contrast discrimination in either of these chromatic pathways, however there is reason to believe they may differ from the achromatic pathway. At detection threshold, binocular summation is greater for chromatic versus achromatic stimuli (Simmons, 2005), implying a more linear initial stage of processing. For cross-orientation masking, there are differences in the magnitude of masking between chromatic (red/green) and achromatic stimuli (Kim et al., 2013; Medina \& Mullen, 2009), as well as differences in their temporal dynamics (Kim \& Mullen, 2015). There are also interactions between chromatic and achromatic pathways both within (Chen et al., 2000) and between (Kingdom \& Libenson, 2015; Mullen et al., 2014) the eyes, yet these have not been fully explored for arrangements where the target and mask have the same orientation. Finally, the neurophysiological underpinnings of colour vision are distinct from the achromatic system. In primary visual cortex (V1), chromatic signals are processed in `blob' regions that are revealed by cytochrome oxidase staining (Horton \& Hubel, 1981). The blob regions appear to be largely monocular (Livingstone \& Hubel, 1984), suggesting that binocular combination for chromatic stimuli might occur later than for achromatic stimuli, and perhaps be subject to different constraints.

@@

All experiments took place in a darkened room. Participants placed their heads in a chin rest mounted on a height-adjustable table, to which the stereoscope was also attached. The total optical viewing distance (including the light path through the mirrors) was 104cm, at which distance 1 degree of visual angle encompassed 48 pixels on the monitor.

Before beginning primary data collection, each participant in Experiments 1 and 2 completed an isoluminance adjustment task. Stimuli were presented that counterphase flickered at 5Hz, defined about either the L-M or S-(L+M) plane in cone space. Participants used a trackball to dynamically adjust the colour angle of the stimulus to minimise the percept of flicker. Each participant completed ten such trials for each colour plane, and the average angle across repetition was taken as the isoluminant point, and used to generate stimuli for the main experiment for that participant. Settings were very similar across participants for the S-(L+M) direction, and somewhat more heterogeneous for the L-M direction (see Figure \ref{fig:isofig}).

@@

In Experiment 1, participants completed a two-interval-forced-choice (2IFC) contrast discrimination task. Stimuli were presented for 200ms, with an interstimulus interval of 400ms. Each interval was indicated by an auditory beep, and participants made their responses using a two-button trackball. Correct trials were indicated by a high pitched tone, and incorrect trials by a low pitched tone. Each block of the experiment tested a single pedestal contrast level, and lasted around 12 minutes. On each trial the target contrast level was determined by a 3-down-1-up staircase procedure. There were 8 interleaved staircases in total; four stimulus arrangements (see Figure \ref{fig:exampledips}) by two target eye assignments. Each pedestal contrast was repeated 3 times by each participant, and the block order was randomised. The experiment lasted around 4 hours per participant for each chromatic condition, and took place over the course of several weeks. In total, the experiment consisted of 67978 trials (pooled across participants).

@@

In Experiment 3, the achromatic conditions from Experiment 1 were repeated using a flickering disc stimulus. The stimulus counterphase flickered at 4Hz, and was presented for 500ms (i.e.~2 full cycles of the temporal modulation). All other procedures were the same as for Experiment 1, and the experiment comprised a total of 24610 trials (pooled across participants).

@@

\begin{equation}
Stage1_L = \frac{C_L^m}{S + C_L + \omega C_R},
\end{equation}

\begin{equation}
Stage1_R = \frac{C_L^m}{S + C_R + \omega C_L},
\end{equation}

\begin{equation}
binsum = Stage1_L + Stage1_R,
\end{equation}

\begin{equation}
binresp = \frac{binsum^p}{Z + binsum^q},
\end{equation}

where \(C_L\) and \(C_R\) are the contrasts displayed to the left and right eyes, and \(m\), \(S\), \(\omega\), \(p\), \(q\) and \(Z\) are free parameters in the model. A further free parameter, \(k\), represents additive internal noise, and is used to convert the model outputs to either d-prime or threshold values. Thresholds are defined by adjusting the target contrast until the following equality is satisfied:

\begin{equation}
binresp_{target+pedestal} - binresp_{pedestal} = k,
\end{equation}

and d-prime for a single target level is defined as:

\begin{equation}
d' = \frac{binresp_{target+pedestal} - binresp_{pedestal}}{k}.
\end{equation}

We ran the simplex algorithm from 100 random starting vectors for each data set, and chose the solution for each data set that gave the smallest RMS error.

We also implemented a Bayesian hierarchical version of the model using the Stan language (Carpenter et al., 2017). This used a binomial generator function to model the proportion correct data at each target level, and was fit simultaneously to all participants for a given experiment, but separately for each chromatic condition of Experiment 1, and the flickering disc data from Experiment 3 (i.e.~four fits in total, as for the simplex fitting). This modelling primarily focuses on examining posterior parameter distributions, rather than a model comparison approach. We generated over 1 million posterior samples for the model, and retained 10\% of them for plotting.

@@

All experimental code, raw data and analysis scripts are available at: \url{https://osf.io/3vdga/}. The linked GitHub repository also contains a fully reproducible version of the manuscript. Note that we deviated slightly from the planned preregistration, in that we did not collect data for chromatic flickering discs, or for the cross-pathway dichoptic experiment using disc stimuli. This is because the grating data from Experiments 1 and 2, and the achromatic disc data from Experiment 3, were sufficient to address the questions we had hoped to answer from these experiments.

@@

Dipper functions from Experiment 1 are displayed in the upper row of Figure \ref{fig:dipperfig}. Panel (a) shows the achromatic results, which replicate the key features from previous work. At detection threshold, binocular summation was a factor of 1.67 (4.47dB), consistent with previous reports (Baker et al., 2018). Pedestal masking functions follow the typical `dipper' shape in all conditions, with a region of facilitation at low pedestal contrasts, and masking at higher contrasts. The monocular and binocular dipper handles converge at high contrasts, whereas the half-binocular thresholds remain above the binocular thresholds across the full range of pedestal contrasts. The dichoptic condition produced very high thresholds, with the rising portion of the dipper having a slope around 1 (regression slope of 1.06 in log (dB) units, calculated across the highest 4 pedestal contrasts).

@@

\caption{Dipper functions and psychometric slopes from Experiment 1, averaged across three participants. Panels (a-c) show threshold data, and panels (d-f) represent the slope of the psychometric function expressed in Weibull beta units. Error bars give ±1SE across participants. Note that contrast values are expressed as a percentage of the maximum displayable contrast (see Procedures for details). Curves in panels (a-c) show the best fitting models, optimized using a simplex algorithm (described in section 4.4), and RMSE values give the root mean square errors of the fits.}\label{fig:dipperfig}
\end{figure}

A similar pattern of results was observed for both the red/green and blue/yellow isoluminant stimuli (see Figure \ref{fig:dipperfig}b,c). Summation at threshold was a factor of 1.71 (4.66dB) for the red/green targets, and a factor of 1.73 (4.77dB) for the blue/yellow targets, and so was marginally higher than for achromatic stimuli. The general character of the dipper functions was largely consistent with the achromatic results, though we observed shallower facilitation and weaker masking, especially for the blue/yellow stimuli. For example, the strongest facilitation in the binocular condition for achromatic stimuli was a factor of 2.63, whereas it reduced to a factor of 2.35 for red/green stimuli and 1.57 for blue/yellow stimuli. The slope of the binocular dipper handle was 0.52 for achromatic stimuli, 0.58 for red/green stimuli, and 0.34 for blue/yellow stimuli. Dichoptic masking remained as strong for the chromatic conditions as for the achromatic stimuli (regression slopes of 1.3 for red/green and 1.21 for blue/yellow). The pattern of results for individual participants was consistent with the group averages, as shown in Figure \ref{fig:individualdips}.

Following Meese et al. (2006), we also inspected the slope of the psychometric function for each condition (see Figure \ref{fig:dipperfig}d-f). At detection threshold, slopes were relatively steep, with \(\beta\) values around 4. As pedestal contrasts increased, slopes linearized and reduced to around \(\beta=1.3\) (Foley \& Legge, 1981; Meese et al., 2006), and remained shallow at high pedestal contrasts. The exception to this was the dichoptic condition, where slopes became extremely steep at high dichoptic mask contrasts, consistent with previous observations (Baker et al., 2013; Meese et al., 2006). This was clear for all three data sets, with slope values in the range \(\beta=4\) to \(\beta=8\). However, we note that slope estimates are more variable than threshold estimates (note the large error bars), particularly when using adaptive staircases, which deploy the majority of trials close to threshold. Our second experiment investigated the slope of the psychometric function in more detail using the method of constant stimuli.

@@

In Experiment 2 we focussed on the dichoptic condition at a single mask contrast, and measured full psychometric functions using the method of constant stimuli for all factorial pairings of target and mask chromaticity. The pooled results across three participants are shown in Figure \ref{fig:MCSfig}a-c, and results for individual participants are available in Figure \ref{fig:individualMCS}. All conditions produced monotonically increasing psychometric functions (panels a-c), but the extent of masking was highly dependent on the relationship between the target and mask chromaticity. Figure \ref{fig:MCSfig}d-f shows a two-dimensional representation of individual threshold and slope estimates (points), as well as the posterior density estimates for fits to the pooled data (ellipses). The results are consistent between participants, and at the group level, and show that the presence of a mask has a strong effect on thresholds.

@@

We also calculated the slope of the psychometric function for each condition, in equivalent Weibull \(\beta\) units. In the absence of a mask, the average slope was \(\beta\) = 3.42, which is typical for contrast detection tasks. Slopes became substantially steeper when the dichoptic mask matched the target in chromaticity (average \(\beta\) = 5.38). These `super-steep' psychometric functions for dichoptic pedestal masking have been reported previously (Baker et al., 2013; Meese et al., 2006), and are observed for the first time here using chromatic stimuli (see diagonal values in Figure \ref{fig:MCSfig}h, and also Figure \ref{fig:dipperfig}d-f). However we did not see such markedly steep functions for any of the cross-chromaticity masking conditions (average \(\beta\) = 2.83 for the off-diagonal values).

@@

In our final experiment, we again measured dipper functions, but this time for a temporally modulating luminance disc. This was motivated by our recent work (Segala et al., 2023) that appeared to show increased binocular facilitation and reduced interocular suppression for flickering disc stimuli (relative to gratings), measured using EEG and a psychophysical matching paradigm. The pattern of dipper functions for a 4Hz flickering disc (see Figure \ref{fig:discdata}a) was very similar to that observed for achromatic gratings (see Figure \ref{fig:dipperfig}a), and the binocular summation ratio at threshold was also similar (a factor of 1.49). Threshold data for individual participants are shown in Figure \ref{fig:individualdiscs}. We also found a similar pattern of psychometric slope values (Figure \ref{fig:discdata}b), though we note that the dichoptic condition did not produce the `super-steep' psychometric functions we had observed in Experiments 1 \& 2. Nevertheless, its slopes are somewhat above those of the other pedestal arrangements.

{\centering \includegraphics[width=0.8\linewidth]{Figures/discdata}

\caption{Thresholds for the flickering disc experiment. Plotting conventions mirror those in Figure 3.}\label{fig:discdata}
\end{figure}

\hypertarget{computational-modelling}{%
\subsection{Computational modelling}\label{computational-modelling}}

\begin{table}
